Simulating High-Dimensional Multivariate Data using the Bigsimr R Package

### 2021-10-19 arXiv_bigsimr/arXiv_bigsimr.tex
Working on cross references arXiv

We illustrate our methodology using the Breast Invasive Carcinoma (BRCA)
data set housed in The Cancer Genome Atlas (TCGA; see Acknowledgments).
For ease of modeling and simplicity of exposition, we only consider high
expressing genes. In turn, we begin by filtering to retain the top r d
of the highest-expressing genes (in terms of median expression) of the
over 20,000 gene measurements from \(N=r nrow(example_brca)\) patients'
tumor samples. This gives a great number of pairwise dependencies among
the marginals (specifically, \(r choose(d,2)\) correlation parameters).
Table @ref(tab:ch010-realDataTab) displays RNA-seq counts for three
selected high-expressing genes for the first five patients' breast tumor
samples. To help visualize the bivariate relationships for these three
selected genes across all patients, Figure @ref(fig:ch010-realDataFig)
displays the marginal distributions and estimated Spearman's
correlations.

@@

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch010-realDataFig-1.pdf}
\caption{Pairwise scatterplots for three example genes (Table 1), with
estimated Spearman's correlations and marginal density estimates
displayed. The data possess outliers, heavy-right tails, are discrete,
and have non-trivial intergene correlations. Modeling these data
motivate our simulation methodology.}
\end{figure}

@@

To place \texttt{Bigsimr}'s implementation and algorithms in context of
existing approaches, we qualitatively compare \texttt{Bbigsimr} to other
multivariate simulation packages. In \texttt{R}, the \texttt{MASS}
package's \texttt{mvnorm()} has long been available to produce random
vectors from multivariate normal (and \(t\)) distributions. This
procedure cannot simulate from arbitrary margins and dependency measures
and it fails to readily scale to high dimensions. SAS PROC SIMNORMAL and
Python's NumPy \texttt{random.multivariate\_normal} both do similar
tasks. Recently, the high-performance \texttt{R} package
\texttt{mvnfast} (Fasiolo 2016) has been released, which provides HD
multivariate normal generation. However, lacks flexibility in marginal
models and dependency modeling. For dependent discrete data, LD
\texttt{R} packages exist, such as \texttt{GenOrd}. For copula
computation, the full-featured \texttt{R} package \texttt{copula}
provides many different copulas and a fairly general LD random generator
\texttt{rCopula()}. High-dimesional random vector generation via
\texttt{copula} takes an impractical amount of computing time (X. Li et
al. 2019). The \texttt{nortaRA} R package matches Pearson coefficients
near exactly and is reasonable for LD settings. Finally, we know of no
other \texttt{R} package that can produce HD random vectors with
flexible margins and dependence structure. Table @ref(tab:compare-table)
summarizes the above discussion.

@@

This section describes our methods for simulating a random vector
\(\bf Y\) with \(Y_i\) components for \(i=1,\ldots,d\). Each \(Y_i\) has
a specified marginal CDF \(F_i\) and its inverse \(F^{-1}_i\). To
characterize dependency, every pair \((Y_i, Y_j)\) has a given Pearson
correlation \(\rho_P\), Spearman correlation \(\rho_S\), and/or
Kendall's \(\tau\). The method can be described as a
\emph{high-performance Gaussian copula} (Equation @ref(eq:gauss))
providing a HD NORTA-inspired algorithm.

@@

\emph{Step 1: Mapping step.} We first employ the closed-form
relationships between \(\rho_S\) and \(\tau\) with \(\rho_P\) for
bivariate normal random variables via Equations @ref(eq:convertKendall)
and @ref(eq:convertSpearman), respectively (implemented as
\texttt{cor\_covert()}). Initializing our algorithm to match the
nonparametric correlations by computing these equations for all pairs is
computationally trivial.

For the computationally expensive process to match Pearson correlations,
we approximate Equation @ref(eq:pearsonIntegralRelation) for all pairs
of margins. To this end, we implement the approximation scheme
introduced by (Xiao and Zhou 2019). Briefly, the many double integrals
of the form in @ref(eq:pearsonIntegralRelation) are approximated by
weighted sums of Hermite polynomials. Matching coefficients for pairs of
continuous distributions is made tractable by this method, but for
discrete distributions (especially discrete distributions with large
support sets or infinite support), the approximation is computationally
expensive. To remedy this, we further approximate discrete distributions
by a continuous distribution using Generalized S-Distributions (Muino,
Voit, and Sorribas 2006).

@@

There is an increasing probability of encountering an inadmissible
correlation matrix as dimension increases. In our experience, the
mapping step for large \(d\) almost always produces a \(R_X\) that is
not a correlation matrix. In Section @ref(package), we provide a basic
method of how to quantify and control the approximation error. Further,
the RNA-seq example in Section @ref(examples) provides an illustration
of this in practice.

@@

This section describes a bivariate random vector simulation workflow via
the \texttt{bigsimr} \texttt{R} package. This \texttt{R} package
provides an interface to the native code written in Julia (registered as
the \texttt{Bigsimr} Julia package). In addition to the native Julia
\texttt{Bigsimr} package and \texttt{R} interface \texttt{bigsimr}, we
also provide a Python interface \texttt{bigsimr} that interfaces with
the Julia \texttt{Bigsimr} package. The Julia package provides a
high-performance implementation of our proposed random vector generation
algorithm and associated functions (see Section @ref(algorithms)).

@@

Figure @ref(fig:ch030-aq-joint-dist) visualizes the bivariate
relationship between \texttt{Ozone} and \texttt{Temperature}. We aim to
simulate random two-component vectors mimicking this structure. The
margins are not normally distributed; \texttt{Ozone} level exhibits a
strong positive skew.

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch030-aq-joint-dist-1.pdf}
\caption{Bivariate scatterplot of Ozone versus Temp with estimated
marginal densities. The Ozone data are modeled marginally as log-normal
and the Temperature data as normal.}
\end{figure}

@@

Based on the estimated densities in Figure
@ref(fig:ch030-aq-joint-dist), we assume \texttt{Temp} is normally
distributed and \texttt{Ozone} is log-normally distributed, as the
latter values are positive and skewed. We use the classical unbiased
estimators for the normal distribution's parameters and maximum
likelihood estimators for the log-normal parameters:

@@

\begin{verbatim}
## Julia Object of type NamedTuple{(:lower, :upper),Tuple{Float64,Float64}}.
## (lower = -0.8816136274594583, upper = 0.8815214379668451)
\end{verbatim}

@@

Finally, we execute \texttt{rvec} to simulate the desired \(10,000\)
random vectors from the assumed bivariate distribution of \texttt{Ozone}
and \texttt{Temp}. Note that the target input is specified using the
pre-computed MVN correlation matrix. Figure @ref(fig:ch030-plot-sim)
plots the 10,000 simulated points.

@@

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch030-plot-sim-1.pdf}
\caption{Contour plot and marginal densities for the simulated bivariate
distribution of Air Quality Temperatures and Ozone levels. The simulated
points mimic the observed data with respect to both the marginal
characteristics and bivariate association.}
\end{figure}

@@

Table @ref(tab:sims) lists our identical-marginal, bivariate simulation
configurations. We increase the simulation replicates \(B\) to visually
ensure that our results converge to the target correlations and gauge
efficiency. We select distributions beginning with a standard
multivariate normal (MVN) as we expect the performance to be exact (up
to MC error) for all correlation types. Then, we select a non-symmetric
continuous distribution: a standard (rate =1), two-component
multivariate gamma (MVG). Finally, we select distributions and marginal
parameter values that are motivated by our RNA-seq data, namely values
proximal to probabilities and sizes estimated from the data (see
\href{examples}{RNA-seq data application} for estimation details).
Specifically, we assume a MVNB with
\(p_1 = p_2 = 3\times10^{-4}, r_1 = r_2 = 4, \rho \in \Omega\).

@@

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-biNormPlot-1.pdf}
\caption{\texttt{bigsimr} recovers the Pearson specified correlations
for MVN.}
\end{figure}

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-biGammaPlot-1.pdf}
\caption{\texttt{bigsimr} recovers the Pearson specified correlations
for Bivariate Gamma.}
\end{figure}

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-biNegBinPlot-1.pdf}
\caption{\texttt{bigsimr} recovers the correlations for bivariate
negative binomial only approximately for Pearson but (nearly) exactly
for the rank-based correlations.}
\end{figure}

Figure @ref(fig:ch040-bPlot) displays the aggregated bivariate
simulation results. Table @ref(tab:ch040-BiError) contains the mean
absolute error (MAE) in reproducing the desired dependency measures for
the three bivariate scenarios. Overall, the studies show that our
methodology is generally accurate across the entire range of possible
correlation for all three dependency measures. Our Pearson matching
performs nearly as well as Spearman or Kendall, except for a slight
increase in error for the negative binomial case.

@@

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-bPlot-1.pdf}
\caption{Bivariate simulations match target correlations across the
entire range of feasible correlations. The horizontal axis plots the
specified target correlations for each bivariate margin. Normal margins
are plotted in dark dark grey, gamma in medium grey, and negative
binomial in light grey. As the number of simulated vectors \(B\)
increases from left to right, the variation in estimated correlations
(vertical axis) decreases. The dashed line indicates equality between
the specified and estimated correlations.}
\end{figure}

@@

The experiments are conducted on a MacBook Pro carrying a 2.4 GHz 8-Core
Intel Core i9 processor, with all 16 threads employed during
computation. Table @ref(tab:ch040-moderateDtab) displays the total
computation time for moderate dimesions (\(d \leq 500\)). In every
simullation setting, the 1,000 vectors are generated rapidly, executing
in under 2 seconds.

@@

The results for \(d > 500\) show scalability to ultra-high dimensions
for all three correlation types, although the total times do become much
larger. Figure @ref(fig:ch040-largeDfig) displays computation times for
\(d=\{1000, 2500, 5000, 10000\}\). For \(d\) equal to 1000 and 2500, the
total time is under a couple of minutes. At \(d\) of 5000 and 10,000,
Pearson correlation matching in the \emph{Adjust Correlation} step
becomes costly. Interestingly, Pearson is actually faster than Kendall
for \(d=10,000\) due to bottlenecks in \emph{Compute Correlation} and
\emph{Check Admissibility}. Uniformly, matching Spearman correlations is
faster, with total times under 5 minutes for \(d=10,000\), making
Spearman the most computationally-friendly dependency type for our
package. With this in mind, we scaled the simulation to \(d=20,000\) for
the Spearman type and obtained the 1,000 vectors in under an hour (data
not shown). In principle, this would enable the simulation of an entire
human-derived RNA-seq data set. We note that for a given target
correlation matrix and margins, steps 1, 2, and 3 only need to be
computed once and the fourth step, \emph{Simulate Data}, is nearly
instaneous for all settings considered.

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-largeDfig-1.pdf}
\caption{Computation times for HD multivariate gamma simulation.}
\end{figure}

@@

Figure @ref(fig:ch050-simDataFig) displays the simulated counts and
pairwise relationships for our example genes from Table
@ref(tab:ch010-realDataTab). Simulated counts roughly mimic the observed
data but with a smoother appearance due to the assumed parametric form
and with less extreme points than the observed data in Figure
@ref(fig:ch010-realDataFig). Figure @ref(fig:ch050-figBRCA) compares the
specified target parameter (horizontal axis) with the corresponding
quantities estimated from the simulated data (vertical axis). The
evaluation shows that the simulated counts approximately match the
target parameters and exhibit the full range of estimated correlations
from the data.

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch050-simDataFig-1.pdf}
\caption{Simulated data for three selected high-expressing genes
generally replicates the estimated data structure. The data do not
exhibit outlying points, but do possess the desired Spearman
correlations, central tendencies, and discrete values.}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\linewidth]{../fig/ch050-figBRCA} \caption{Simulated random vectors from a multivariate negative binomial replicate the estimated structure from an RNA-seq data set. The dashed red lines indicate equality between estimated parameters from simulated data (vertical axes) and the specified target parameters (horizontal axes).}\label{fig:ch050-figBRCA}
\end{figure}

>>>>

We illustrate our methodology using the Breast Invasive Carcinoma (BRCA)
data set housed in The Cancer Genome Atlas (TCGA; see Acknowledgments).
For ease of modeling and simplicity of exposition, we only consider high
expressing genes. In turn, we begin by filtering to retain the top r d
of the highest-expressing genes (in terms of median expression) of the
over 20,000 gene measurements from \(N=r nrow(example_brca)\) patients'
tumor samples. This gives a great number of pairwise dependencies among
the marginals (specifically, \(r choose(d,2)\) correlation parameters).
Table \ref{tab:ch010-realDataTab} displays RNA-seq counts for three
selected high-expressing genes for the first five patients' breast tumor
samples. To help visualize the bivariate relationships for these three
selected genes across all patients, Figure \ref{fig:ch010-realDataFig}
displays the marginal distributions and estimated Spearman's
correlations.

@@

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch010-realDataFig-1.pdf}
\caption{\label{fig:ch010-realDataFig}Pairwise scatterplots for three
example genes (Table 1), with estimated Spearman's correlations and
marginal density estimates displayed. The data possess outliers,
heavy-right tails, are discrete, and have non-trivial intergene
correlations. Modeling these data motivate our simulation methodology.}
\end{figure}

@@

To place \texttt{Bigsimr}'s implementation and algorithms in context of
existing approaches, we qualitatively compare \texttt{Bbigsimr} to other
multivariate simulation packages. In \texttt{R}, the \texttt{MASS}
package's \texttt{mvnorm()} has long been available to produce random
vectors from multivariate normal (and \(t\)) distributions. This
procedure cannot simulate from arbitrary margins and dependency measures
and it fails to readily scale to high dimensions. SAS PROC SIMNORMAL and
Python's NumPy \texttt{random.multivariate\_normal} both do similar
tasks. Recently, the high-performance \texttt{R} package
\texttt{mvnfast} (Fasiolo 2016) has been released, which provides HD
multivariate normal generation. However, lacks flexibility in marginal
models and dependency modeling. For dependent discrete data, LD
\texttt{R} packages exist, such as \texttt{GenOrd}. For copula
computation, the full-featured \texttt{R} package \texttt{copula}
provides many different copulas and a fairly general LD random generator
\texttt{rCopula()}. High-dimesional random vector generation via
\texttt{copula} takes an impractical amount of computing time (X. Li et
al. 2019). The \texttt{nortaRA} R package matches Pearson coefficients
near exactly and is reasonable for LD settings. Finally, we know of no
other \texttt{R} package that can produce HD random vectors with
flexible margins and dependence structure. Table \ref{tab:compare-table}
summarizes the above discussion.

@@

This section describes our methods for simulating a random vector
\(\bf Y\) with \(Y_i\) components for \(i=1,\ldots,d\). Each \(Y_i\) has
a specified marginal CDF \(F_i\) and its inverse \(F^{-1}_i\). To
characterize dependency, every pair \((Y_i, Y_j)\) has a given Pearson
correlation \(\rho_P\), Spearman correlation \(\rho_S\), and/or
Kendall's \(\tau\). The method can be described as a
\emph{high-performance Gaussian copula} (Equation \ref{eq:gauss})
providing a HD NORTA-inspired algorithm.

@@

\emph{Step 1: Mapping step.} We first employ the closed-form
relationships between \(\rho_S\) and \(\tau\) with \(\rho_P\) for
bivariate normal random variables via Equations \ref{eq:convertKendall}
and \ref{eq:convertSpearman}, respectively (implemented as
\texttt{cor\_covert()}). Initializing our algorithm to match the
nonparametric correlations by computing these equations for all pairs is
computationally trivial.

For the computationally expensive process to match Pearson correlations,
we approximate Equation \ref{eq:pearsonIntegralRelation} for all pairs
of margins. To this end, we implement the approximation scheme
introduced by (Xiao and Zhou 2019). Briefly, the many double integrals
of the form in \ref{eq:pearsonIntegralRelation} are approximated by
weighted sums of Hermite polynomials. Matching coefficients for pairs of
continuous distributions is made tractable by this method, but for
discrete distributions (especially discrete distributions with large
support sets or infinite support), the approximation is computationally
expensive. To remedy this, we further approximate discrete distributions
by a continuous distribution using Generalized S-Distributions (Muino,
Voit, and Sorribas 2006).

@@

There is an increasing probability of encountering an inadmissible
correlation matrix as dimension increases. In our experience, the
mapping step for large \(d\) almost always produces a \(R_X\) that is
not a correlation matrix. In Section \ref{package}, we provide a basic
method of how to quantify and control the approximation error. Further,
the RNA-seq example in Section \ref{examples} provides an illustration
of this in practice.

@@

This section describes a bivariate random vector simulation workflow via
the \texttt{bigsimr} \texttt{R} package. This \texttt{R} package
provides an interface to the native code written in Julia (registered as
the \texttt{Bigsimr} Julia package). In addition to the native Julia
\texttt{Bigsimr} package and \texttt{R} interface \texttt{bigsimr}, we
also provide a Python interface \texttt{bigsimr} that interfaces with
the Julia \texttt{Bigsimr} package. The Julia package provides a
high-performance implementation of our proposed random vector generation
algorithm and associated functions (see Section \ref{algorithms}).

@@

Figure \ref{fig:ch030-aq-joint-dist} visualizes the bivariate
relationship between \texttt{Ozone} and \texttt{Temperature}. We aim to
simulate random two-component vectors mimicking this structure. The
margins are not normally distributed; \texttt{Ozone} level exhibits a
strong positive skew.

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch030-aq-joint-dist-1.pdf}
\caption{\label{fig:ch030-aq-joint-dist}Bivariate scatterplot of Ozone
versus Temp with estimated marginal densities. The Ozone data are
modeled marginally as log-normal and the Temperature data as normal.}
\end{figure}

@@

Based on the estimated densities in Figure
\ref{fig:ch030-aq-joint-dist}, we assume \texttt{Temp} is normally
distributed and \texttt{Ozone} is log-normally distributed, as the
latter values are positive and skewed. We use the classical unbiased
estimators for the normal distribution's parameters and maximum
likelihood estimators for the log-normal parameters:

@@

\begin{verbatim}
## Julia Object of type NamedTuple{(:lower, :upper),Tuple{Float64,Float64}}.
## (lower = -0.8807315594247829, upper = 0.880286004475508)
\end{verbatim}

@@

Finally, we execute \texttt{rvec} to simulate the desired \(10,000\)
random vectors from the assumed bivariate distribution of \texttt{Ozone}
and \texttt{Temp}. Note that the target input is specified using the
pre-computed MVN correlation matrix. Figure \ref{fig:ch030-plot-sim}
plots the 10,000 simulated points.

@@

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch030-plot-sim-1.pdf}
\caption{\label{fig:ch030-plot-sim}Contour plot and marginal densities
for the simulated bivariate distribution of Air Quality Temperatures and
Ozone levels. The simulated points mimic the observed data with respect
to both the marginal characteristics and bivariate association.}
\end{figure}

@@

Table \ref{tab:sims} lists our identical-marginal, bivariate simulation
configurations. We increase the simulation replicates \(B\) to visually
ensure that our results converge to the target correlations and gauge
efficiency. We select distributions beginning with a standard
multivariate normal (MVN) as we expect the performance to be exact (up
to MC error) for all correlation types. Then, we select a non-symmetric
continuous distribution: a standard (rate =1), two-component
multivariate gamma (MVG). Finally, we select distributions and marginal
parameter values that are motivated by our RNA-seq data, namely values
proximal to probabilities and sizes estimated from the data (see
\href{examples}{RNA-seq data application} for estimation details).
Specifically, we assume a MVNB with
\(p_1 = p_2 = 3\times10^{-4}, r_1 = r_2 = 4, \rho \in \Omega\).

@@

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-biNormPlot-1.pdf}
\caption{\label{fig:ch040-biNormPlot}\texttt{bigsimr} recovers the
Pearson specified correlations for MVN.}
\end{figure}

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-biGammaPlot-1.pdf}
\caption{\label{fig:ch040-biGammaPlot}\texttt{bigsimr} recovers the
Pearson specified correlations for Bivariate Gamma.}
\end{figure}

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-biNegBinPlot-1.pdf}
\caption{\label{fig:ch040-biNegBinPlot}\texttt{bigsimr} recovers the
correlations for bivariate negative binomial only approximately for
Pearson but (nearly) exactly for the rank-based correlations.}
\end{figure}

Figure \ref{fig:ch040-bPlot} displays the aggregated bivariate
simulation results. Table \ref{tab:ch040-BiError} contains the mean
absolute error (MAE) in reproducing the desired dependency measures for
the three bivariate scenarios. Overall, the studies show that our
methodology is generally accurate across the entire range of possible
correlation for all three dependency measures. Our Pearson matching
performs nearly as well as Spearman or Kendall, except for a slight
increase in error for the negative binomial case.

@@

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-bPlot-1.pdf}
\caption{\label{fig:ch040-bPlot}Bivariate simulations match target
correlations across the entire range of feasible correlations. The
horizontal axis plots the specified target correlations for each
bivariate margin. Normal margins are plotted in dark dark grey, gamma in
medium grey, and negative binomial in light grey. As the number of
simulated vectors \(B\) increases from left to right, the variation in
estimated correlations (vertical axis) decreases. The dashed line
indicates equality between the specified and estimated correlations.}
\end{figure}

@@

The experiments are conducted on a MacBook Pro carrying a 2.4 GHz 8-Core
Intel Core i9 processor, with all 16 threads employed during
computation. Table \ref{tab:ch040-moderateDtab} displays the total
computation time for moderate dimesions (\(d \leq 500\)). In every
simullation setting, the 1,000 vectors are generated rapidly, executing
in under 2 seconds.

@@

The results for \(d > 500\) show scalability to ultra-high dimensions
for all three correlation types, although the total times do become much
larger. Figure \ref{fig:ch040-largeDfig} displays computation times for
\(d=\{1000, 2500, 5000, 10000\}\). For \(d\) equal to 1000 and 2500, the
total time is under a couple of minutes. At \(d\) of 5000 and 10,000,
Pearson correlation matching in the \emph{Adjust Correlation} step
becomes costly. Interestingly, Pearson is actually faster than Kendall
for \(d=10,000\) due to bottlenecks in \emph{Compute Correlation} and
\emph{Check Admissibility}. Uniformly, matching Spearman correlations is
faster, with total times under 5 minutes for \(d=10,000\), making
Spearman the most computationally-friendly dependency type for our
package. With this in mind, we scaled the simulation to \(d=20,000\) for
the Spearman type and obtained the 1,000 vectors in under an hour (data
not shown). In principle, this would enable the simulation of an entire
human-derived RNA-seq data set. We note that for a given target
correlation matrix and margins, steps 1, 2, and 3 only need to be
computed once and the fourth step, \emph{Simulate Data}, is nearly
instaneous for all settings considered.

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch040-largeDfig-1.pdf}
\caption{\label{fig:ch040-largeDfig}Computation times for HD
multivariate gamma simulation.}
\end{figure}

@@

Figure \ref{fig:ch050-simDataFig} displays the simulated counts and
pairwise relationships for our example genes from Table
\ref{tab:ch010-realDataTab}. Simulated counts roughly mimic the observed
data but with a smoother appearance due to the assumed parametric form
and with less extreme points than the observed data in Figure
\ref{fig:ch010-realDataFig}. Figure \ref{fig:ch050-figBRCA} compares the
specified target parameter (horizontal axis) with the corresponding
quantities estimated from the simulated data (vertical axis). The
evaluation shows that the simulated counts approximately match the
target parameters and exhibit the full range of estimated correlations
from the data.

\begin{figure}
\centering
\includegraphics{arXiv_bigsimr_files/figure-latex/ch050-simDataFig-1.pdf}
\caption{\label{fig:ch050-simDataFig}Simulated data for three selected
high-expressing genes generally replicates the estimated data structure.
The data do not exhibit outlying points, but do possess the desired
Spearman correlations, central tendencies, and discrete values.}
\end{figure}

\begin{figure}
\includegraphics[width=0.8\linewidth]{../fig/ch050-figBRCA} \caption{\label{fig:ch050-figBRCA}Simulated random vectors from a multivariate negative binomial replicate the estimated structure from an RNA-seq data set. The dashed red lines indicate equality between estimated parameters from simulated data (vertical axes) and the specified target parameters (horizontal axes).}\label{fig:ch050-figBRCA}
\end{figure}
